Towards a Practical Face Recognition System: \\ Robust Alignment and Illumination by Sparse Representation

### 2011-03-31 rebuttal.tex

   \title{\Large {\bf Manuscript  Revision Statement}}

   \author{{Andrew Wagner, John Wright, Arvind Ganesh, Zihan Zhou, Hossein Mobahi, and Yi Ma}}

   \date{}
   \maketitle

We again thank the reviewers and the area chair for their comments.
%I am mostly satisfied with the responses to the earlier review. The paper is
%in much better form now and more carefully written.
Changes have been made to the paper to address all of the changes requested by Reviewer 1,
including a more complete set of results in Table 2, Table 3, Figure 11, and associated
discussions.  We give a quick overview of the changes.

@@

{\bf
	In any case, the extent of manual input is unclear. It is said 'the input face
	is aligned to the training with manually selected outer eye corners'. If only
	2 points are manually selected, then what kind of registration can be done
	with just 2 corresponding points ? Surely not affine alignment. Many more are
	needed if it is to be done robustly. So it is not surprising that manual
	alignment is giving worse results than automatic. If my interpretation is
	wrong, is the entire affine registration being done manually ? If so, as
	mentioned above the accuracy of the human can be bad.
}

Similarity transformations are used for all of the experiments in this section, and
therefore two manually selected outer eye corners are sufficient.  We have
made this clear by adding a footnote.

{\bf
	Some comments on this would be useful to better understand the results shown in table 1 and 2.
}

The discussion of these results (seen in Table 2 and 3 in the new
draft) has been updated. For the single gallery image FERET
experiment (Table 3), our algorithm is not designed for
this scenario, and it fails when the illumination changes and other
variations seriously violate its assumptions (this was true in the previous
revision as well). We found that LBP
with manual alignment achieves the best performance in this
experiment.

### 2011-03-31 faces_system.tex
Fixed spacing

\subsection{Recognition with Pose and Expression} We now run tests of
our algorithm on a subset of the images from Multi-PIE with pose and expression variation in the test set, although we do not model these variations explicitly.
Using the same training set as above, we test our algorithm on
images in Session 2 with $15^\circ$ pose, for all 20
illuminations. As expected, the recognition rate drops to 78.0\%. We also test our
algorithm on images in Session 3 with smile, again for all 20
illuminations. The recognition rate is 64.8\%. Of course, it is reasonable to expect that
the performance of our method will be significantly improved if pose and expression data
are available in the training.

\section{Tests on Our Own Database}\label{sec:own-data} Using the training acquisition
system we described in Section \ref{sec:illumination}, and shown in Figure
\ref{fig:system}, we have collected the frontal view of 109
subjects {\em without eyeglasses} under 38 illuminations shown
in Figure \ref{fig:illumination-patterns}. For testing our
algorithm, we have also taken 935 images of these subjects with
a different camera under a variety of practical conditions.

\subsection{Necessity of Rear Illuminations} To see how
training illuminations affect the performance of our algorithm
in practice, we now compare how well a few frontal
illuminations can linearly represent: 1. other frontal illuminations
taken under the same laboratory conditions, and 2. typical
indoor and outdoor illuminations. To this end, we use the face
database acquired by our system and use 7 illuminations per
subject as training. The illuminations are chosen to be similar
to the 7 illuminations used in the previous experiment on
Multi-PIE.\footnote{We use the illuminations $\{6, 9, 12,
13, 18, 21, 22\}$ shown in Figure
\ref{fig:illumination-patterns}(b) to mimic the illuminations
 $\{0, 1, 6, 7, 13, 14, 18\}$ in Multi-PIE.} We then test
our algorithm on the remaining $24 - 7 = 17$ frontal
illuminations for all the subjects. The recognition rate is
$99.8\%$, nearly perfect. We also test our algorithm on 310
indoor images and 168 outdoor images of these subjects taken
under a variety of lighting conditions (category 1 and 2
specified below), similar to the one shown in Figure
\ref{fig:promo}, and the recognition rates for indoor and
outdoor images drop down to $94.2\%$ and $89.2\%$,
respectively. This is a strong indication that
frontal illuminations taken under laboratory conditions
are insufficient for representing test images under typical indoor and
outdoor illuminations.\vspace{-0.2in}

@@

\begin{figure}[t]
\centering
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_1565.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3656.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3827.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4090.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4106.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4126.JPG} \\
\vspace{2mm}
\centering
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_1611.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3528.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3744.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3995.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4030.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4095.JPG}
\vspace{-.0in}
 \caption{{\bf Representative examples of category C4}. Top row: successful examples with our method using overlapping blocks. Bottom row: failures with our method using overlapping blocks.}\label{fig:examples4}
\vspace{-.5in}
\end{figure}

@@

\begin{figure}
\centerline{
\includegraphics[width=4in]{figures_pami/occ_block.png}\vspace{-0.2in}}
\caption{{\bf Using overlapping blocks to tackle contiguous occlusion.} (a) The test image, occluded by sunglasses. (b) The four overlapping blocks. (c) The sparse representation is calculated after alignment for each block independently. The red lines correspond to his true identity. (d) The true identity is successfully recovered by voting based on the SCI scores.}
\label{fig:occ-block}
\vspace{-.5in}
\end{figure}

>>>>

\vspace{-.2in}
\subsection{Recognition with Pose and Expression} We now run tests of
our algorithm on a subset of the images from Multi-PIE with pose and expression variation in the test set, although we do not model these variations explicitly.
Using the same training set as above, we test our algorithm on
images in Session 2 with $15^\circ$ pose, for all 20
illuminations. As expected, the recognition rate drops to 78.0\%. We also test our
algorithm on images in Session 3 with smile, again for all 20
illuminations. The recognition rate is 64.8\%. Of course, it is reasonable to expect that
the performance of our method will be significantly improved if pose and expression data
are available in the training.

\vspace{-.1in}
\section{Tests on Our Own Database}\label{sec:own-data} Using the training acquisition
system we described in Section \ref{sec:illumination}, and shown in Figure
\ref{fig:system}, we have collected the frontal view of 109
subjects {\em without eyeglasses} under 38 illuminations shown
in Figure \ref{fig:illumination-patterns}. For testing our
algorithm, we have also taken 935 images of these subjects with
a different camera under a variety of practical conditions.

\vspace{-.1in}
\subsection{Necessity of Rear Illuminations} To see how
training illuminations affect the performance of our algorithm
in practice, we now compare how well a few frontal
illuminations can linearly represent: 1. other frontal illuminations
taken under the same laboratory conditions, and 2. typical
indoor and outdoor illuminations. To this end, we use the face
database acquired by our system and use 7 illuminations per
subject as training. The illuminations are chosen to be similar
to the 7 illuminations used in the previous experiment on
Multi-PIE.\footnote{We use the illuminations $\{6, 9, 12,
13, 18, 21, 22\}$ shown in Figure
\ref{fig:illumination-patterns}(b) to mimic the illuminations
 $\{0, 1, 6, 7, 13, 14, 18\}$ in Multi-PIE.} We then test
our algorithm on the remaining $24 - 7 = 17$ frontal
illuminations for all the subjects. The recognition rate is
$99.8\%$, nearly perfect. We also test our algorithm on 310
indoor images and 168 outdoor images of these subjects taken
under a variety of lighting conditions (category 1 and 2
specified below), similar to the one shown in Figure
\ref{fig:promo}, and the recognition rates for indoor and
outdoor images drop down to $94.2\%$ and $89.2\%$,
respectively. This is a strong indication that
frontal illuminations taken under laboratory conditions
are insufficient for representing test images under typical indoor and
outdoor illuminations.
\vspace{-0.1in}

@@

\begin{figure}[t]
\centering
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_1565.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3656.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3827.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4090.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4106.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4126.JPG} \\
\vspace{2mm}
\centering
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_1611.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3528.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3744.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3995.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4030.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4095.JPG}
\vspace{-.0in}
 \caption{{\bf Representative examples of category C4}. Top row: successful examples with our method using overlapping blocks. Bottom row: failures with our method using overlapping blocks.}\label{fig:examples4}
\vspace{-.4in}
\end{figure}

@@

\begin{figure}
\centerline{
\includegraphics[width=4in]{figures_pami/occ_block.png}\vspace{-0.2in}}
\caption{{\bf Using overlapping blocks to tackle contiguous occlusion.} (a) The test image, occluded by sunglasses. (b) The four overlapping blocks. (c) The sparse representation is calculated after alignment for each block independently. The red lines correspond to his true identity. (d) The true identity is successfully recovered by voting based on the SCI scores.}
\label{fig:occ-block}
\vspace{-.3in}
\end{figure}
